---
title: "PM566-lab3"
output:
  html_document:
    toc: yes
    toc_float: yes
    fig_height: 10
    fig_width: 10
    #code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading the Data
```{r message = FALSE}
library(tidyverse)
if (!require(data.table)) {
  install.packages("data.table")
}
if (!file.exists("met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}
met <- data.table::fread("met_all.gz")
```

# Check the dimensions
```{r message = FALSE}
dim(met)
```
# Check Header
```{r message = FALSE}
head(met)
```
# Check Footers
```{r message = FALSE}
tail(met)
```
## There are 2,377,343 rows and 30 columns in the met dataset.

# Looking at the Variables 
```{r message = FALSE}
str(met)
```
# Key Variables

```{r message = FALSE}
table(met$year)
table(met$day)
table(met$hour)
```

# Looking at Continuous Variables 
```{r message = FALSE}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

# Updating NAs in elevation
```{r message = FALSE}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

The highest weather station is at `r max(met$elev,na.rm = TRUE)` ft.

# Remove Observations that have Min Temperature of -40C
```{r message = FALSE}
met <- met[temp>-40][order(temp)]
summary(met)
```
# Removed Temperatures colder than -15C, making the new dataset have a minimum temp of -3C. 
```{r message = FALSE}
met <- met[temp>-15][order(temp)]
summary(met[, .(lat, lon, wind.sp, temp, elev)])
```
# Calculating Summary Statistics 
```{r message = FALSE}
met[elev==max(elev,na.rm=TRUE)][, summary(wind.sp)]
met[elev==max(elev,na.rm=TRUE)][, summary(temp)]
```
# Correlations
```{r message = FALSE}
met[elev ==max(elev, na.rm=TRUE), .(
  temp_wind= cor(temp, wind.sp, use ="complete"),
  temp_day=cor(temp, wind.sp, use ="complete"),
  temp_hour=cor(temp, wind.sp, use ="complete"),
  wind_day=cor(temp, wind.sp, use ="complete"),
  wind_hour=cor(temp, wind.sp, use ="complete")
)]
```

# Exploratory Graphs 
```{r message = FALSE}
hist(met$elev, breaks=100)
```

```{r message = FALSE }
hist(met$temp)
```

```{r message = FALSE}
hist(met$wind.sp)
```

```{r message = FALSE}
library(leaflet)
elev <- met[elev==max(elev,na.rm=TRUE)]
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r message = FALSE}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
elev <- elev[order(date)]
head(elev)
```

```{r message = FALSE}
plot(elev$date, elev$temp, type='l')
```
## The temperature seems to drop around the end of August. 
```{r message = FALSE}
plot(elev$date, elev$wind.sp, type='l')
```
## The Windspeed seems to spike around mid August and the end of August. 