---
title: "Lab_11"
output: 
  html_document:
    toc: yes
    toc_float: yes
    html_preview: false
  github_document:
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSQLite)
library(DBI)
```

# Setup
```{r}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")
# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")
# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
```

```{r}
dbListTables(con)
```
# Exercise 1: Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.
```{r}
dbGetQuery(con,
      "SELECT actor_id, last_name, first_name
      FROM actor
      ORDER by last_name, first_name
      LIMIT 5")
```

# Excercise 2: Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.
```{r}
dbGetQuery(con,
      "SELECT actor_id, last_name, first_name
      FROM actor
      WHERE last_name IN ('WILLIAMS', 'DAVIS') 
      LIMIT 5")
```

# Exercise 3: Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.
```{r}
dbGetQuery(con,
      "SELECT DISTINCT customer_id
      FROM rental
      WHERE date(rental_date) = '2005-07-05'
      LIMIT 5")
```

# Exercise 4
## Exercise 4.1: Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{r}
q <- dbSendQuery(con,"
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)")
dbFetch(q, n=10)
```

## Exercise 4.2: Construct a query that retrives all rows from the payment table where the amount is greater then 5
```{r}
q <- dbSendQuery(con,"
SELECT *
FROM payment
WHERE amount > 5")
dbFetch(q, n=10)
```

## Exercise 4.3: Construct a query that retrives all rows from the payment table where the amount is greater then 5 and less then 8
```{r}
q <- dbSendQuery(con,"
SELECT *
FROM payment
WHERE amount > 5 and amount < 8")
dbFetch(q, n=10)
```

## Exercise 5: Retrive all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.
```{r}
dbGetQuery(con,"
SELECT p.payment_id, p.amount
FROM payment AS p
  INNER JOIN customer AS c ON p.customer_id=c.customer_id
where c.last_name = 'DAVIS'"
)
```

# Exercise 6
## Exercise 6.1: Use COUNT(*) to count the number of rows in rental
```{r}
dbGetQuery(con,"
SELECT COUNT (*)
FROM rental
")

```

## Exercise 6.2: Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id
```{r}
dbGetQuery(con,"
SELECT customer_id, COUNT (*) AS 'N Rentals'
FROM rental GROUP BY customer_id
LIMIT 5
")
```

## Exercise 6.3: Repeat the previous query and sort by the count in descending order
```{r}
dbGetQuery(con,"
SELECT customer_id, COUNT (*) AS 'N Rentals'
FROM rental GROUP BY customer_id
ORDER BY COUNT(*) DESC
LIMIT 5
")
```

## Exercise 6.4: Repeat the previous query but use HAVING to only keep the groups with 40 or more.
```{r}
dbGetQuery(con,"
SELECT customer_id, COUNT (*) AS 'N Rentals'
FROM rental GROUP BY customer_id
HAVING 'N Rentals' >= 40
ORDER BY COUNT(*) DESC
LIMIT 5
")
```

# Exercise 7:
```{r}
dbGetQuery(con,"
SELECT MAX(amount) AS 'max', 
MIN(amount) AS 'min', 
AVG(amount) AS 'avg', 
SUM(amount) AS 'sum'
FROM payment
")
```


## Exercise 7.1: Modify the above query to do those calculations for each customer_id
```{r}
dbGetQuery(con,"
SELECT
  customer_id,
  MAX(amount) AS 'max', 
  MIN(amount) AS 'min', 
  AVG(amount) AS 'avg', 
  SUM(amount) AS 'sum'
FROM payment GROUP BY customer_id
LIMIT 5
")
```

## Excerise 7.2: Modify the above query to only keep the customer_ids that have more then 5 payments
```{r}
dbGetQuery(con,"
SELECT
  customer_id,
  MAX(amount) AS 'max', 
  MIN(amount) AS 'min', 
  AVG(amount) AS 'avg', 
  SUM(amount) AS 'sum'
FROM payment 
GROUP BY customer_id
HAVING COUNT(*) >5 
LIMIT 5
")
```

# Cleanup 
```{r}
# clean up
dbDisconnect(con)
```

