---
title: "LAB_5"
author: "Ram Ayyala"
date: "9/24/2021"
output: 
  github_document:
    html_preview: false
  html_document:
    toc: yes
    toc_float: yes
    fig_height: 10
    fig_width: 10
    #code_folding: hide
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries and data 
```{r}
library(data.table)
library(tidyverse)
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]
```
#Stations Data Processing 
```{r}
# Dealing with NAs and 999999
#stations[stations$USAF==999999] <- NA_integer_
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

# Met Data Processing
```{r message=FALSE}
met <- met[temp>=-17][order(temp)]
summary(met)

##Make sure there are no missing data in the key variables coded as 9999, 999, etc
met[met$elev==9999.0] <- NA
summary(met$elev)
```

# Merging Data
```{r}
dat <- merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )
```

# Question 1: Representative station for the US
## What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the quantile() function. Do these three coincide?

We generate a representative version of each station. We will use th eaverages (median could also be a good way to represent it, but ti will depend on case)
```{r collapsing-by-station}
station_averages <- dat[,.(
  temp = mean(temp,na.rm=TRUE),
  wind.sp = mean(wind.sp, na.rm=TRUE),
  atm.press = mean(atm.press, na.rm=TRUE)
), by= USAFID]
```

# Identify Quantiles per variable
```{r}
medians <- station_averages[,.(
  temp_50 = quantile(temp,probs = .5, na.rm=TRUE),
  wind.sp_50 = quantile(wind.sp,probs = .5, na.rm=TRUE),
  atm.press_50 = quantile(atm.press,probs = .5, na.rm=TRUE)
)]
medians
```
Now we can find stations that match these values or are the closets to them via `which.min()` 
```{r}
station_averages[, temp_dist := abs(temp -medians$temp_50)]
median_temp_station <- station_averages[order(temp_dist)][1]
median_temp_station
```
The Median temperature station is ` r median_temp station$USAFID`. 

# Question 2: Representative station per state 
## Just like the previous question, you are asked to identify what is the most representative, the median, station per state. This time, instead of looking at one variable at a time, look at the euclidean distance. If multiple stations show in the median, select the one located at the lowest latitude.
```{r}
station_averages <- merge(
  # Data
  x     = station_averages,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )
station_averages[, temp_50 := quantile(temp, probs=.5, na.rm=TRUE), by= STATE]
station_averages[, wind.sp_50 := quantile(wind.sp, probs=.5, na.rm=TRUE), by= STATE]
station_averages

#euclidean distance
station_averages[, eudist := sqrt((temp-temp_50)^2 + (wind.sp - wind.sp_50)^2)]
```


# Question 3: In the middle? 
## For each state, identify what is the station that is closest to the mid-point of the state. Combining these with the stations you identified in the previous question, use leaflet() to visualize all ~100 points in the same figure, applying different colors for those identified in this question.
```{r}

```

# Question 4: Means of means
## Using the quantile() function, generate a summary table that shows the number of states included, average temperature, wind-speed, and atmospheric pressure by the variable “average temperature level,” which you’ll need to create.

Start by computing the states’ average temperature. Use that measurement to classify them according to the following criteria:

low: temp < 20
Mid: temp >= 20 and temp < 25
High: temp >= 25
Once you are done with that, you can compute the following:

Number of entries (records),
Number of NA entries,
Number of stations,
Number of states included, and
Mean temperature, wind-speed, and atmospheric pressure.
All by the levels described before.
```{r}
#avg temp by state 
dat[, state_temp := mean(temp, na.rm=TRUE), by = STATE]
#Create temp cat variable
dat[, temp_cat := fifelse(
  state_temp < 20, "low-temp",
  fifelse(state_temp < 25, "mid-temp", "high-temp"))
  ]
#Check for NAs
table(met$temp_cat, useNA="always")

# Summarize
tab <-dat [, .(
  N_entries =.N,
  N_stations = length(unique(USAFID)),
  N_states = length(unique(STATE)),
  temp_avg= mean(temp, na.rm=TRUE),
  wind.sp_avg = mean(wind.sp, na.rm=TRUE),
  atm.press_avg = mean(atm.press, na.rm=TRUE)
), by=temp_cat]

knitr::kable(tab)

```





