---
title: "Supplementary"
author: "Ram Ayyala"
date: "12/10/2021"
output: 
  html_document:
    toc: yes
    toc_float: yes
    html_preview: false
link-citations: yes
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
if (!require(tidyverse)) {
  install.packages("tidyverse")} 
if (!require(data.table)) {
  install.packages("data.table")}
if (!require(leaflet)) {
  install.packages("leaflet")}
if (!require(hablar)) {
  install.packages("hablar")}
if (!require(Amelia)) {
  install.packages("Amelia")}
if (!require(ggpubr)) {
  install.packages("ggpubr")}
if (!require(forcats)) {
  install.packages("forcats")}
lb <- fread("data/heartdisease_long_beach_va.csv")
sw <- fread("data/heartdisease_switzerland.csv")
hg <- fread("data/heartdisease_hungarian.csv")
cv <- fread("data/heartdisease_cleveland.csv")
```




```{r include=FALSE}
lb <- lb %>%
  mutate(location = "Long Beach")
cv <- cv %>%
  mutate(location = "Cleveland")
hg <- hg %>%
  mutate(location = "Hungary")
sw <- sw %>%
  mutate(location = "Switzerland")
```



```{r include=FALSE}
lb <- lb %>% 
  convert(num(V1,V2,V3,V4,V5,V6,V7,V8,V9,V10,V11),
          chr(V12,V13,location),
          int(V14))
hg <- hg %>% 
  convert(num(V1,V2,V3,V4,V5,V6,V7,V8,V9,V10,V11),
          chr(V12,V13,location),
          int(V14))
sw <- sw %>% 
  convert(num(V1,V2,V3,V4,V5,V6,V7,V8,V9,V10,V11),
          chr(V12,V13,location),
          int(V14))
```



```{r include=FALSE}
lb <- lb %>% 
  map_df(.,.f=function(x){
             ifelse(as.character(x) == "?",
             NA,as.character(x))})
cv <- cv %>% 
  map_df(.,.f=function(x){
             ifelse(as.character(x) == "?",
             NA,as.character(x))})
hg <- hg %>% 
  map_df(.,.f=function(x){
             ifelse(as.character(x) == "?",
             NA,as.character(x))})

sw <- sw %>% 
  map_df(.,.f=function(x){
             ifelse(as.character(x) == "?",
             NA,as.character(x))})

```



```{r include=FALSE}
data <- merge(merge(merge(
  cv,
  hg, all = TRUE),
  sw, all = TRUE),
  lb, all = TRUE)
```


```{r include=FALSE}
# Reconvert columns to correct data type
data <- data %>% 
  convert(num(V1,V2,V3,V4,V5,V6,V7,V8,V9,V10,V11,V12,V13),
          chr(location),
          int(V14))

colnames(data) <- c('age','sex','cp','trestbps','cholserum','fbs','restecg','thalach','exang','oldpeakST','slope','ca','thal','num','location')

data <- as.data.frame(data)
```


```{r include=FALSE}
idvars= c(names(which(!colSums(is.na(data)) > 0)))
c(names(which(colSums(is.na(data)) > 0)))
a.out <- amelia(data, m = 5, idvars=idvars)
data <- a.out$imputations[[1]]
data <- data.table(data)

data$fbs[data$fbs < 0.5]  <- 0
data$fbs[data$fbs >=0.5]  <- 1

data$restecg[data$restecg < 0.5]  <- 0
data$restecg[data$restecg >=0.5]  <- 1
data$restecg[data$restecg <1.5]  <- 1
data$restecg[data$restecg >=1.5]  <- 2
data$restecg[data$restecg >=2]  <- 2

data$exang[data$exang < 0.5]  <- 0
data$exang[data$exang >=0.5]  <- 1

data$slope[data$slope < 0.5]  <- 0
data$slope[data$slope >=0.5]  <- 1
data$slope[data$slope <1.5]  <- 1
data$slope[data$slope >=1.5]  <- 2
data$slope[data$slope >=2]  <- 2

data$ca[data$ca < 0.5]  <- 0
data$ca[data$ca >=0.5]  <- 1
data$ca[data$ca <1.5]  <- 1
data$ca[data$ca >=1.5]  <- 2
data$ca[data$ca < 2.5]  <- 2
data$ca[data$ca >=2.5]  <- 3
data$ca[data$ca <3.5]  <- 3
data$ca[data$ca >=3.5]  <- 4
data$ca[data$ca >=4]  <- 4

data$thal[data$thal < 0.5]  <- 0
data$thal[data$thal >=0.5]  <- 1
data$thal[data$thal <1.5]  <- 1
data$thal[data$thal >=1.5]  <- 2
data$thal[data$thal < 2.5]  <- 2
data$thal[data$thal >=2.5]  <- 3
data$thal[data$thal >=3]  <- 3


```


```{r include=FALSE}
data <- data[cholserum>0][order(cholserum)]
summary(data$cholserum)
```

 
```{r include=FALSE}
data <- data[data$num == "0" | data$num == "1", ]
data[num=="0", sick := "< 50% diameter narrowing"]
data[num=="1", sick := "> 50% diameter narrowing"]

data[, .(
  number_obv = .N
), by = sick] %>%
  knitr::kable()

```



<!--# Create cat variable cholesterol level -->

```{r include=FALSE}

# > 240 = high
# 200 to 239 = borderline high
# <200 is desirable 
# based on US Department of Health and Human Services, Mayo Clinic and Cleveland Clinic Standards

data[cholserum < 200, cholesterol_level := "Desirable"]
data[cholserum >= 200 & cholserum < 240, cholesterol_level := "Borderline High"]
data[cholserum >= 240, cholesterol_level := "High"]

t1<-data[, .(
  min_cholserum = min(cholserum),
  max_cholserum = max(cholserum),
  number_obv = .N
), by = cholesterol_level] %>%
  knitr::kable(col.names= c("Cholesterol Level", "Minimum Cholesterol Level (mg/dl)", "Maximum Cholesterol Level (mg/dl)", "Number of Observations"), caption="Cholesterol Level Categories")

```



```{r include=FALSE}
dim(data)
```

```{r include=FALSE}
head(data) %>% knitr::kable()
tail(data) %>% knitr::kable()
```

```{r include=FALSE}
str(data)
```


```{r include=FALSE}
summary(data$age)
summary(data$sex)
summary(data$cholserum)
summary(data$sick)
#Checking for missing values
table(is.na(data$age))
table(is.na(data$sex))
table(is.na(data$cholserum))
table(is.na(data$sick))
```
```{r include=FALSE}
s1<-data[!is.na(location)] %>%
  ggplot(mapping=aes(x=cholserum, y=trestbps, color=location)) +
  geom_point() +
  geom_smooth(method="lm", color="black") +
  facet_wrap(~ location) +
  labs(title = "Cholesterol Levels vs Resting Blood Pressure", x= "Cholesterol Level (mg/dl)", y="Resting blood pressure (mmHg)") +
   scale_color_brewer(palette = "Spectral")
```


```{r include=FALSE}
s2<-data[!is.na(location) & !is.na(cholesterol_level)] %>%
  mutate(location=fct_reorder(location, cholserum, na.rm=TRUE)) %>%
   mutate(cholesterol_level=fct_reorder(cholesterol_level, cholserum, na.rm=TRUE)) %>%
  ggplot() +
  geom_bar(mapping=aes(x=location, fill =cholesterol_level), position="dodge")+
  labs(title = "Location by Cholesterol Level", x= "Location", y="Count") +
  scale_fill_brewer(palette = "Spectral")
```



## Supplementary Figures {.tabset}

### Supplementary Figure 1

```{r echo=FALSE}
s1
```
 <br />
The figure above contains a scatter plot that details the relationship between Cholesterol Levels and Resting blood pressure for each location in the data set, with the x-axis being the Cholesterol Levels (mg/dl) and the y-axis being the Resting Blood pressure values (mmHg). Within each scatter plot lies a regression line. From the figure above, it is evident that across all of the locations, Cholesterol seems to have a positive association with Resting Blood Pressure.

### Supplementary Figure 2

```{r echo=FALSE}
s2
```
<br />
The graph above depicts a bar chart of Location by Cholesterol Level, with the x-axis being the Location from which the samples were obtained from and the y-axis being the count. From the graph, we can see that the majority of each location is made up of individuals with high cholesterol, while a very small portion have the desirable level of cholesterol. Moreover, for each location, the order of the cholesterol level category in terms of count proportion seems to be maintained in this order from largest to smallest: High, Borderline High, and Desirable. 
{-}