---
title: "Assignment 4"
author: "Ram Ayyala"
date: "11/19/2021"
output: 
  github_document: 
  html_document:
    toc: yes
    toc_float: yes
    html_preview: false
  word_document: 
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(microbenchmark)
library(parallel)
```

# HPC
## Problem 1: Make sure your code is nice
### Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google
```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  t(apply(mat,1,cumsum))
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), "relative","equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), "relative", "equivalent"
)
```

## Problem 2: Make things run faster with parallel computing
### The following function allows simulating PI
```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

```{r}
system.time({
  cl <- makePSOCKcluster(4L)
  clusterSetRNGStream(cl, 1231)
  clusterExport(cl,c("sim_pi"), envir=environment())
  ans <- unlist(parLapply(cl=cl,1:4000,sim_pi, n=10000))
  print(mean(ans))
  stopCluster(cl)
})
```

#SQL
## Setup a temporary database by running the following chunk
```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

## Question 1:
### How many many movies is there avaliable in each rating catagory.
```{sql, connection=con}

SELECT rating, 
COUNT (*) AS "Film Number"
FROM film
GROUP BY rating

```
From the Table above, we see how many movies are available for each category. 

## Question 2:
### What is the average replacement cost and rental rate for each rating category.
```{sql, connection=con}
SELECT rating,
AVG(replacement_cost) AS "Average Replacement Cost",
AVG(rental_rate) AS "Average Rental Rate"
FROM film
GROUP BY rating
```
From the Table above, we see that the average replacement cost and average rental rate for each rating category. It is evident that the PG-13 movies seem to have the highest avergae replacement cost while PG movies have the highest average rental rate. Moreover, it seems that PG movies have the lowest average replacement cost while G movies have the lowest average rental rate. 

## Question 3:
Use table film_category together with film to find the how many films there are with each category ID
```{sql, connection=con}
SELECT category_id, 
COUNT(title) as "Number of Films"
FROM film
INNER JOIN film_category
on film.film_id = film_category.film_id
GROUP BY category_id
```
From the table above, we see that **category_id 15** has the most number of films at 74 films while **category_id 12** has the least number of films at 51 films. 

## Question 4:
Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql, connection=con}
SELECT name, 
COUNT(title) as "Number of Films"
FROM film
INNER JOIN film_category
on film.film_id = film_category.film_id
INNER JOIN category
on category.category_id = film_category.category_id
GROUP BY name
ORDER BY -"Number of Films" 
```
From the above table, we see that the most popular category is the Sports category with 74 films. 

```{r}
dbDisconnect(con)
```

